---
title: "awu3_hw3"
format: html
editor: visual
---

## Task 1: Conceptual Questions

-   <div>

    1.  **If your working directory is myfolder/homework/, what relative path would you specify to get the file located at myfolder/MyData.csv?**
        -   The relative path to get to the file located at myfolder/MyData.csv would be ../MyData.csv as ../ would take it back a directory to myfolder where the file is located
    2.  **What are the major benefits of using R projects?**
        -   R projects provide a way to divide work into multiple contexts with each having a working directory, workspace, history, and folder structure.
    3.  **What is Git and what is Github?**
        -   Git is a version control software that helps track changes in source code over time. While Github is an online hosting service for Git-based projects so many people can access the same source code.
    4.  **What are the two main differences between a tibble and a data.frame?**
        -   A tibble will have a special way of printing out the contents(data) which is different than a data.frame also when sub setting to 1 column using \[\] they do not coerce down to a vector.

    </div>

------------------------------------------------------------------------

## Task 2: Reading Delimited Data

**1.)** Read this data into R directly from the URL using functions from the tidyverse. Notice that the data doesn’t include column names - add those (in a manner of your choosing). Print out the tibble (just call the object name)

```{r}
# import tidyverse package
library(tidyverse)

# read glass dataset and create df
glass_df <- read_delim( "https://www4.stat.ncsu.edu/online/datasets/glass.data", delim=",", col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K","Ca","Ba","Fe", "Glass_type"))

glass_df

```

**2.)** Read this data (Yeast Data) into R directly from the URL using functions from the tidyverse. Notice that the data doesn’t include column names - add those (in a manner of your choosing). Print out the tibble (just call the object name).

```{r}

# read yeast data and create df
yeast_df <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data", col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class"))

yeast_df
```

## Task 3: Combing Excel and Delimited Data

**3.)** There is an excel version of the white wine data set available at https://www4.stat.ncsu.edu/\~online/datasets/white-wine.xlsx. • Download this file • Place it in a folder you know (such as your working directory for your project, the folder your .qmd file lives in, a data folder you’ve created in your repo, etc.) • Import the data from the first sheet using the readxl package • Print out the tibble (just call the object name)

```{r}
# import readxl package
library(readxl)

# read in white wine csv and create df
white_wine_df <- read_excel("data/white-wine.xlsx")

white_wine_df
```

**4.)**When you print the data set out to the console, you may notice that some of the variable names are surrounded by backticks. This is because they are non-standard (they include a space in them). We can rename them in a number of ways. We’ll do it by reading in the variable names from the 2nd sheet of the same file.

• Read in the data from the 2nd sheet. This should return a data frame with one column containing alternative versions of the variable names.

• Grab that column and overwrite the current column names (colnames()) of your white wine tibble. Note that colnames(object) allows for ‘left hand side assignment’. That is, you can call it on the left and store an object in it to overwrite it. Similar to names(object) \<- c(...)

```{r}

# get col names from the second sheet in white wine dataset
wine_col_names <- read_xlsx("data/white-wine.xlsx", sheet = excel_sheets("data/white-wine.xlsx")[2])

wine_col_names

# set new col names from the second sheet
colnames(white_wine_df) <- wine_col_names$Variables

white_wine_df
```

**5.)**Lastly, add a column to this data set to indicate the wines are white. That is, add a column that hasvalues of ‘white’ for every observation. You can do this via the \$ operator or, if you know it already, the mutate() function from dplyr.

```{r}

# label all white wines with type col "white"
white_wine_df$type <- "white"

white_wine_df
```

**6.)** There is a semi-colon delimited version of the red wine data set available at <https://www4.stat.ncsu.> edu/\~online/datasets/red-wine.csv.

• Read this in using the readr package. Be careful that the columns are read in as the correct type!

• You should replace the variable names as done above

• You should append a column denoting the type as “red”

```{r}

# read in red wine csv 
red_wine_df <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", delim= ";")

# set col names with col names from the csv
colnames(red_wine_df) <- wine_col_names$Variables
red_wine_df$type <- "red"


red_wine_df
```

**7.)**. Combine these two data sets into one data set. They both have the exact same columns so this is an easy append task!

• Use the dplyr::bind_rows() function (see the help) to create one tibble containing all of the wine data.

```{r}

# combine white wine and red wine datasets
full_wine_df <- bind_rows(white_wine_df, red_wine_df)

full_wine_df
```

## Task 4: Writing a Function

**8.)**Write a basic function (call it getIQR) that takes in a vector of numbers and returns the interquartile range (75th percentile - 25th percentile). You can use the quantile function. Remove the name on the returned vector (i.e. make sure the returned item doesn’t have a name like 75% or something like that). Allow your function to pass unnamed arguments to the quantile function

```{r}
# function to get IQR range
getIQR <- function(x, ...) {
   quantile(x, probs = 0.75, names= FALSE, ...) - quantile(x, probs = .25, names= FALSE, ...)
  
}
  
```

**9.)** Take the object from question 7 and apply the getIRQ() function to the fixed_acidity and citric_acid columns (two function calls).

```{r}

# call the function on fixed_acifity and citric_acid cols
getIQR(full_wine_df$fixed_acidity)
getIQR(full_wine_df$citric_acid)
```

**10.)**Write another function (call it getCoefVar) that returns the sample coefficient of variation, defined as the sample standard deviation divided by the absolute value of the sample mean. Allow the function to pass unnamed arguments to sd and mean.

```{r}
# function that returns the sample coefficient of variation
getCoefVar <- function(x,...) {
  sd(x,...) / mean(x, ...)
}
```

**11)** Repeat question 9 using the getCoefVar() function

```{r}
# call the getCoefVar for fixed_acidity and citric_acid cols
getCoefVar(full_wine_df$fixed_acidity)
getCoefVar(full_wine_df$citric_acid)
```

**12.)** Create a function (call it getSummaries) that takes in a vector and returns both the IQR and the coefficient of variation. Do not rewrite your above two functions, call them. When returning your values, give them appropriate names. (This would be called a wrapper function.)

```{r}
# function that gets both the IQR and getCoefVar and returns both
getSummaries <- function(x) {
  cat("IQR: " , getIQR(x), "\n")
  cat("Coef Variance: " , getCoefVar(x), "\n")
}
```

**13)**Repeat question 2 with the getSummaries() function.

```{r}
# call the getSummaries function on fixed_acidity and citric_acid cols
getSummaries(full_wine_df$fixed_acidity)
getSummaries(full_wine_df$citric_acid)
```

**14)**Write your own infix function that can be used in replace of the cbind function (feel free to call thecbind function in your function :) Check that the function works on the two data frames below (df1 & df2) and then all three objects (df1, df2, vec1)!

```{r}
# create data frames for the infix function
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)

# infix function that calls cbind
`%cbind%` <- function(x,y) {
  cbind(x,y)
}

df1 %cbind% df2

df1 %cbind% df2 %cbind% vec1


```
