---
title: "emma_hw_3"
format: html
editor: visual
---

# TASK ONE

1.  If your working directory is myfolder/homework/, what relative path would you specify to get the file located at myfolder/MyData.csv?
    1.  read.csv('MyData.csv')
2.  What are the major benefits of using R projects?
    1.  An R project folder can be used as a working directory, so the paths to things should work for other users.
    2.  Other people can download an R project and work on it seamlessly, allowing for easy collaboration.
3.  What is Git and what is GitHub?
    1.  Git is a type of version control software.
    2.  Github is a hosting service for Git based projects.
4.  What are the two main differences between a tibble and a data.frame?
    1.  The default print() method for a tibble is different from a data.frame.
    2.  Tibbles do not coerce down to a vector when you subset to only one column using \[.

# TASK TWO

*Installing tidyverse*

```{r}
install.packages("tidyverse")
```

*Reading in glass and yeast datasets and renaming columns*

```{r}
#Reading in tidyverse
library(tidyverse)

#Reading in glass data and renaming columns
glass_data <- read_csv('https://www4.stat.ncsu.edu/online/datasets/glass.data',
                       col_names = c('Id', 'RI', 'Na', 'Mg', 'Al', 'Si', 'K', 'Ca', 'Ba', 'Fe', 'glass_type'))
glass_data

#Reading in yeast data and renaming columns
yeast_data <- read_table('https://www4.stat.ncsu.edu/online/datasets/yeast.data',
                         col_names = c('gvh', 'alm', 'mit', 'erl', 'pox', 'vac', 'nuc', 'class'))
yeast_data
```

# TASK THREE

*Reading in sheet 1 and 2 for the white wine data, as well as renaming columns and adding in **Color** column. Also reading in red wine data, renaming columns, and adding **Color** column. Finally, both datasets were combined*.

```{r}
#Reading in readxl package
library(readxl)

#Reading in wine data and renaming columns
wine_data <- read_excel('white-wine.xlsx',
                        sheet = excel_sheets('white-wine.xlsx')[1],
                        col_names = c('fixed_acidity', 'volatile_acidity', 'citric_acid', 'residual_sugar', 'chlorides', 'free_sulfur_dioxide', 'total_sulfur_dioxide', 'density', 'pH', 'sulphates', 'alcohol', 'quality'))
wine_data

#Reading in wine data sheet 2
wine_data_2 <- read_excel('white-wine.xlsx',
                          sheet = excel_sheets('white-wine.xlsx')[2])
wine_data_2

#Adding in color column
wine_data$color <- "WHITE"
wine_data

#Reading in red wine data and renaming columns
red_wine_data <- read_csv('red-wine.csv',
                          col_names = c('fixed_acidity', 'volatile_acidity', 'citric_acid', 'residual_sugar', 'chlorides', 'free_sulfur_dioxide', 'total_sulfur_dioxide', 'density', 'pH', 'sulphates', 'alcohol', 'quality'))

#Adding in color column
red_wine_data$color <- 'RED'
red_wine_data

#Combining wine datasets
combined_wine_data <- dplyr::bind_rows(wine_data, red_wine_data, .id = NULL)
combined_wine_data

#Changing fixed acidity and citric acid variables to numeric for future tasks
library(dplyr)
combined_wine_data <- combined_wine_data %>%
  mutate(fixed_acidity = as.numeric(fixed_acidity))
combined_wine_data <- combined_wine_data %>%
  mutate(citric_acid = as.numeric(citric_acid))
```

# TASK FOUR

*Writing **getIQR**, **getCoefVar**, and **getSummaries** functions. Also writing **infix** function and testing to see if it works*.

```{r}
#Writing getIQR function
getIQR <- function(x,...) {
  q <- quantile(x, probs = 0.75, ...) - quantile(x, probs = 0.25, ...)
  unname (q[1]-q[2])
}

#Performing getIQR function on combined wine data variables
getIQR(combined_wine_data$fixed_acidity)
getIQR(combined_wine_data$citric_acid)

#Writing getCoefVar function
getCoefVar <- function(x,...){
  coef <- sd(x,...)/abs(mean(x,...))
}

#Performing getCoefVar function on combined wine data variables
getCoefVar(combined_wine_data$fixed_acidity)
getCoefVar(combined_wine_data$citric_acid)

#Writing getSummaries function
getSummaries <- function(x,...){
  c(IQR = getIQR(x,...), CoefVar = getCoefVar(x,...))
}

#Performing getSummaries function on combined wine data variables
getSummaries(combined_wine_data$fixed_acidity)
getSummaries(combined_wine_data$citric_acid)

#Writing infix function
'%cbind%' <- function(x,y){
  cbind(x,y)
}

#Making data frames and vector
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)

#Testing infix function
df1 %cbind% df2
df1 %cbind% df2 %cbind% vec1
```
