---
title: "emma_hw_3"
format: html
editor: visual
---

# TASK ONE

1.  If your working directory is myfolder/homework/, what relative path would you specify to get the file located at myfolder/MyData.csv?
    1.  read.csv('MyData.csv')
2.  What are the major benefits of using R projects?
    1.  An R project folder can be used as a working directory, so the paths to things should work for other users.
    2.  Other people can download an R project and work on it seamlessly, allowing for easy collaboration.
3.  What is Git and what is GitHub?
    1.  Git is a type of version control software.
    2.  Github is a hosting service for Git based projects.
4.  What are the two main differences between a tibble and a data.frame?
    1.  The default print() method for a tibble is different from a data.frame.
    2.  Tibbles do not coerce down to a vector when you subset to only one column using \[.

# TASK TWO

*Installing tidyverse*

```{r}
install.packages("tidyverse")
```

*Reading in glass and yeast datasets and renaming columns*

```{r}
library(tidyverse)
glass_data <- read_csv('https://www4.stat.ncsu.edu/online/datasets/glass.data',
                       col_names = c('Id', 'RI', 'Na', 'Mg', 'Al', 'Si', 'K', 'Ca', 'Ba', 'Fe', 'glass_type'))
glass_data

yeast_data <- read_table('https://www4.stat.ncsu.edu/online/datasets/yeast.data',
                         col_names = c('gvh', 'alm', 'mit', 'erl', 'pox', 'vac', 'nuc', 'class'))
yeast_data
```

# TASK THREE

```{r}
library(readxl)
wine_data <- read_excel('white-wine.xlsx',
                        sheet = excel_sheets('white-wine.xlsx')[1],
                        col_names = c('fixed_acidity', 'volatile_acidity', 'citric_acid', 'residual_sugar', 'chlorides', 'free_sulfur_dioxide', 'total_sulfur_dioxide', 'density', 'pH', 'sulphates', 'alcohol', 'quality'))
wine_data

wine_data_2 <- read_excel('white-wine.xlsx',
                          sheet = excel_sheets('white-wine.xlsx')[2])
wine_data_2

wine_data$color <- "WHITE"
wine_data

red_wine_data <- read_csv('red-wine.csv',
                          col_names = c('fixed_acidity', 'volatile_acidity', 'citric_acid', 'residual_sugar', 'chlorides', 'free_sulfur_dioxide', 'total_sulfur_dioxide', 'density', 'pH', 'sulphates', 'alcohol', 'quality'))
red_wine_data$color <- 'RED'
red_wine_data

combined_wine_data <- dplyr::bind_rows(wine_data, red_wine_data, .id = NULL)
combined_wine_data
```

# TASK FOUR

```{r}
getIQR <- function(vector){
  return(quantile(0.75) - quantile(0.25))
  col_names = FALSE
}

```
